Using Optimal Time Step Selection to Boost the Accuracy of Finite Difference Methods for a Class of Time Dependent Partial Differential Equations

### 2008-05-14 papers/OTSPDE/OTSPDE.tex
OTSPDE:  minor revisions to captions to make them more consistent.

\begin{figure}[tb]
\begin{center}
\scalebox{0.35}{\includegraphics{figures/diffusion_eqn_1d_no_src_error_vs_N}} 
\caption{$L^\infty$ error as a function of number of grid points for three 
finite difference schemes that solve the 1D diffusion equation: forward Euler 
with optimal time step (circles), forward Euler with suboptimal time step
$\dt = 2\dx^2/3D$ (squares), and Crank-Nicholson with time step 
$\dt = \dx/16 D$ (diamonds).  All three schemes use the standard second-order, 
central finite difference approximation for the Laplacian.  Using each method, 
we solve the diffusion equation with $D = 0.5$ to time $t = 0.1$ on the 
domain $0 < x < 1$ subject to the boundary conditions $u(0,t) = 1$ and 
$u_x(1,t) = 1/2$ and the initial condition
$u(x,0) = 1 + x/2 + 2 \sin(5 \pi x/2) + 3 \sin(7 \pi x/2) 
- 4 \sin(11 \pi x/2)$.  
Note that the errors for the forward Euler with suboptimal time 
step and Crank-Nicholson schemes lie almost directly on top of each other
at the resolution of the figure.
}
\label{fig:diffusion_eqn_1d_no_src_error}
\end{center}
\end{figure}

@@

\begin{figure}[tb]
\begin{center}
\scalebox{0.35}{\includegraphics{figures/diffusion_eqn_1d_src_error_vs_N}} 
\caption{$L^\infty$ error as a function of number of grid points for the 
two finite difference schemes that solve the diffusion equation: 
forward Euler with optimal time step and correction terms (circles) and 
forward Euler with optimal time step but \emph{without} correction terms 
(triangles).  Both schemes use the standard second-order, central finite 
difference approximation for the Laplacian.
Using each method, we solve the diffusion equation with $D = 0.5$ and
source term 
$f(x) = 5 \sin(3 \pi x/2) - 7 \sin(15 \pi x/2) + 10 \sin(21 \pi x/2)$
to time $t = 0.5$ on the domain $0 < x < 1$ subject to the boundary 
conditions $u(0,t) = 1$ and $u_x(1,t) = 1/2$ 
and the initial condition
$u(x,0) = 1 + x/2 + 2 \sin(5 \pi x/2) + 3 \sin(7 \pi x/2) 
- 4 \sin(11 \pi x/2)$.  
}
\label{fig:diffusion_eqn_1d_src_error}
\end{center}
\end{figure}

@@

\begin{figure}[tb]
\begin{center}
\scalebox{0.35}{\includegraphics{figures/adv_diff_eqn_1d_error_vs_N}} 
\caption{$L^\infty$ error as a function of number of grid points for 
two finite difference schemes that solve the advection-diffusion 
equation: forward Euler with optimal time step and correction term (circles) 
and forward Euler with suboptimal time step $\dt = 2\dx^2/3D$ and no correction 
terms (squares).  Both schemes use second-order, central finite difference 
approximations for the advection and diffusion terms.
Using both methods, we solve the advection-diffusion equation with $D = 2$
and $A = 5$ to time $t = 1$ on the domain $-10 < x < 10$ subject to the 
initial condition $u(x,0) = \exp(-x^2/4)$.  The boundary conditions are set 
using the analytical solution 
$u(x,t) = \frac{1}{\sqrt{T}} \exp \left( -(x+At)^2/4T \right)$
where $T = Dt+1$.
}
\label{fig:ADE_1d_error}
\end{center}
\end{figure}

@@

\begin{figure}[tb]
\begin{center}
\scalebox{0.32}{\includegraphics{figures/adv_diff_eqn_1d_FE_OTS_soln}} 
\ \ \ \ \
\scalebox{0.32}{\includegraphics{figures/adv_diff_eqn_1d_FE_soln}} 
\caption{Comparison of numerical solutions for the 1D advection-diffusion
equation~(\ref{eq:ADE_1d}) using forward Euler time stepping with 
optimal time step and correction terms (left) and forward Euler time 
stepping with suboptimal time step $\dt = 2 \dx^2/3 D$ (right).  
Both figures show the numerical solution at $t = 1$ computed on the domain 
$-10 < x < 10$ using 50 grid points.  For details about the initial/boundary 
conditions and physical parameters used to compute these solutions, see the
caption for Figure~\ref{fig:ADE_1d_error}.
}
\label{fig:ADE_1d_solns}
\end{center}
\end{figure}

@@

\subsection{Viscous Burgers Equation}
% FIGURE PLACED HERE TO SPREAD FIGURES OUT
\begin{figure}[tb]
\begin{center}
\scalebox{0.35}{\includegraphics{figures/burgers_eqn_1d_error_vs_N}} 
\caption{$L^\infty$ error as a function of number of grid points for
two finite difference schemes that solve the viscous Burgers equation:
forward Euler with optimal time step and correction term (circles)
and forward Euler with suboptimal time step $\dt = \dx^2/4\nu$ and no 
correction terms (squares).
Both schemes use second-order, central finite difference approximations for
the advection and diffusion terms.
The viscous Burgers equation is solved on the domain $0 < x < 10$ to 
time $t=2$ with the viscosity $\nu$ set to $0.1$.  The initial conditions are 
taken to be 
$u(x,0) = 1 + \gamma \sqrt{\frac{\nu}{\pi}} \exp\left( -(x-1)^2/4 \nu \right) 
            / \left[ 1 + \frac{\gamma}{2} \ 
                         \mathtt{erfc}\left((x-1)/\sqrt{4\nu} \right) 
    \right]$,
where $\ln(1+\gamma)$ is the effective Reynolds number (set equal to 10) and 
the wave speed is set equal to $1$.  The boundary conditions are imposed 
exactly by using
the analytical solution
$u(x,t) = 1 + \gamma \sqrt{\frac{\nu}{\pi T}} 
              \exp\left( -(x-T)^2/4 \nu T \right) 
            / \left[ 1 + \frac{\gamma}{2} \ 
                         \mathtt{erfc}\left((x-T)/\sqrt{4\nu T} \right) 
    \right]$,
where $T = t+1$.
}
\label{fig:burgers_1d_error}
\end{center}
\end{figure}

@@

% FIGURE PLACED HERE TO SPREAD FIGURES OUT
\begin{figure}[tb]
\begin{center}
\scalebox{0.32}{\includegraphics{figures/burgers_eqn_1d_FE_OTS_soln}} 
\ \ \ \ \ 
\scalebox{0.32}{\includegraphics{figures/burgers_eqn_1d_FE_soln}} 
\caption{Comparison of numerical solutions for the 1D viscous Burgers 
equation~(\ref{eq:burgers_1d}) using forward Euler time stepping with 
optimal time step and correction terms (left) and forward Euler time 
stepping with suboptimal time step (right).  Both figures show the 
numerical solution at $t = 2$ computed using $100$ grid points.
For details about the initial/boundary conditions and physical parameters 
used to compute these solutions, see the caption for 
Figure~\ref{fig:burgers_1d_error}. 
}
\label{fig:burgers_1d_solns}
\end{center}
\end{figure}

@@

\begin{figure}[tb]
\begin{center}
\scalebox{0.33}{
  \includegraphics{figures/4th_order_parabolic_eqn_1d_error_vs_N}} 
\ \ 
\scalebox{0.33}{
  \includegraphics{figures/4th_order_parabolic_eqn_1d_comp_time}} 
\caption{$L^\infty$ error as a function of number of grid points (left)
and computation time as a function of $L^\infty$ error (right) for two 
finite difference schemes that solve the fourth-order parabolic equation 
(\ref{eq:4th_order_parabolic_eqn_1d_space_err}):
forward Euler with optimal time step and correction term (circles)
and forward Euler with suboptimal time step $\dt = \dx^2/16$ and no 
correction terms (squares).
Both schemes use a fourth-order, central finite difference approximation for
the bilaplacian.  Using each method, we solve the fourth-order parabolic
equation with $\kappa = 1$ on the domain $0 < x < 1$ subject to the boundary 
conditions $u(0,t) = u(1,t) = 0$ and the initial condition
$u(x,0) = 2\sin(2 \pi x) + \sin(5 \pi x)$.  The value of the solution
in the ghost cells (required as a result of the wide stencil for the 
discrete bilaplacian) are filled by taking values from the exact solution: 
$u(x,t) = 2\sin(2 \pi x)\cos(10000 \pi t) 
        + \sin(5 \pi x) \left[ 2-\exp(-100\pi^2 t) \right]$.
}
\label{fig:4th_order_parabolic_eqn_1d_error}
\end{center}
\end{figure}

@@

\begin{figure}[tb]
\begin{center}
\scalebox{0.33}{\includegraphics{figures/diffusion_eqn_2d_src_error_vs_N}} 
\ \ 
\scalebox{0.33}{\includegraphics{figures/diffusion_eqn_2d_src_comp_time}} 
\caption{$L^\infty$ error as a function of number of grid points (left)
and computation time as a function of $L^\infty$ error (right) for three
finite difference schemes that solve the 2D diffusion equation:
forward Euler with optimal time step and correction terms (circles), forward 
Euler with suboptimal time step (squares), and Crank-Nicholson (diamonds).  
The forward Euler schemes use a nine-point discretization of the Laplacian;
the Crank-Nicholson scheme uses the standard five-point discretization of
the Laplacian.  Using each method, we solve the 2D diffusion equation with 
$D = 0.25$ on the domain $0 < x, y < 1$ subject to the initial condition
$u(x,y,0) =  1 + y/3 + x/2 + xy/4 
          + \sin(2 \pi x) \sin(3 \pi y) - \sin(5 \pi x) \sin(\pi y)$
and boundary conditions taken from the analytical solution 
$u(x,t) = 1 + y/3 + x/2 + xy/4 
          + \sin(2 \pi x) \sin(3 \pi y) \exp(-5 D \pi^2 t) 
          - \sin(5 \pi x) \sin(\pi y) \exp(-2 D \pi^2 t)
$.
Note that the errors for the forward Euler with suboptimal time step and 
Crank-Nicholson schemes lie almost directly on top of each other at the 
resolution of the figures.
}
\label{fig:diffusion_eqn_2d_src_analysis}
\end{center}
\end{figure}

@@

% FIGURE SHOWING SOLUTION OF 2D DIFFUSION EQUATION ON STARFISH DOMAIN
% PLACED HERE TO SPREAD FIGURES OUT
\begin{figure}[tb]
\begin{center}
\scalebox{0.35}{\includegraphics{figures/diffusion_eqn_2d_starfish_domain_soln}} 
\scalebox{0.35}{\includegraphics{figures/diffusion_eqn_2d_starfish_domain_error}} 
\caption{Numerical solutions (left) and dominant error (right) for the 2D 
diffusion equation on a starfish-shaped domain.  The solution is computed at 
time $t = 0.1$ using forward Euler time stepping with an optimal time step 
and correction terms.  In the error plot, the dark regions represent points 
where the error in the solution is larger than $25$\% of the $L^\infty$ error 
of the solution.  For these figures, the diffusion constant $D$ is set to 
$1/4$ and the source term is taken to be 
$f(x,y,t) = 8 D \pi^2 \sin(2 \pi x) \sin(3 \pi y) 
            \exp\left(-5 D \pi^2 t\right)
          - 24 D \pi^2 \sin(5 \pi x) \sin(\pi y) 
            \exp \left(-2 D \pi^2 t \right).
$
The initial conditions for the problem are 
$u(x,y,0) = 1 + y/3 + x/2 + xy/4 
          + \sin(2 \pi x) \sin(3 \pi y) 
          - \sin(5 \pi x) \sin(\pi y)
$
and the boundary conditions are obtained from the analytical solution
$u(x,y,t) = 1 + y/3 + x/2 + xy/4 
              + \sin(2 \pi x) \sin(3 \pi y) \exp(-5 D \pi^2 t) 
              - \sin(5 \pi x) \sin(\pi y) \exp(-2 D \pi^2 t)
$.
The boundary of the starfish domain is represented by the zero-level set of 
the function
$\sqrt{x^2 + y^2} - \frac{3}{5} \left(1 + \frac{1}{2}\sin(5 \theta) \right)$, 
where $\theta = \mathtt{atan}\left( y/x \right)$.
The numerical solution is calculated using the square computational domain 
$-1 < x,y < 1$ which is discretized using 400 grid points in each coordinate 
direction (the plot of the solution has been downsampled for visualization 
purposes).  Boundary conditions are imposed by using fourth-order accurate
approximations of the value of the solution in edge and corner ghost cells
(described in the main text).
}
\label{fig:diffusion_eqn_2d_starfish_domain}
\end{center}
\end{figure}

@@

% FIGURE SHOWING SOLUTION OF 2D DIFFUSION EQUATION ON STARFISH DOMAIN
% PLACED HERE TO SPREAD FIGURES OUT
\begin{figure}[tb]
\begin{center}
\scalebox{0.35}{
  \includegraphics{figures/diffusion_eqn_2d_starfish_domain_error_vs_N}} 
\caption{$L^\infty$ error as a function of number of grid points for two
finite difference schemes that solve the 2D diffusion equation on an 
irregular domain: forward Euler with optimal time step and correction 
terms (circles) and forward Euler with suboptimal time step $\dt = \dx^2/8 D$
(squares).  Both methods solve the 2D diffusion equation described in the 
caption for Figure~\ref{fig:diffusion_eqn_2d_starfish_domain}.
}
\label{fig:diffusion_eqn_2d_starfish_error}
\end{center}
\end{figure}

>>>>

\begin{figure}[tb]
\begin{center}
\scalebox{0.35}{\includegraphics{figures/diffusion_eqn_1d_no_src_error_vs_N}} 
\caption{$L^\infty$ error as a function of number of grid points for three 
finite difference schemes that solve the 1D diffusion equation: forward Euler 
with optimal time step (circles), forward Euler with suboptimal time step
$\dt = 2\dx^2/3D$ (squares), and Crank-Nicholson with time step 
$\dt = \dx/16 D$ (diamonds).  All three schemes use the standard second-order, 
central finite difference approximation for the Laplacian.  Using each method, 
we solve the diffusion equation with $D = 0.5$ to time $t = 0.1$ on the 
domain $0 < x < 1$ subject to the boundary conditions $u(0,t) = 1$ and 
$u_x(1,t) = 1/2$ and the initial condition
$u(x,0) = 1 + x/2 + 2 \sin(5 \pi x/2) + 3 \sin(7 \pi x/2) 
- 4 \sin(11 \pi x/2)$.  
Note that the errors for the forward Euler with suboptimal time 
step and Crank-Nicholson schemes lie almost directly on top of each other
at the resolution of the figure.
These results were obtained using simple MATLAB implementations of the 
finite difference schemes.
}
\label{fig:diffusion_eqn_1d_no_src_error}
\end{center}
\end{figure}

@@

\begin{figure}[tb]
\begin{center}
\scalebox{0.35}{\includegraphics{figures/diffusion_eqn_1d_src_error_vs_N}} 
\caption{$L^\infty$ error as a function of number of grid points for the 
two finite difference schemes that solve the diffusion equation: 
forward Euler with optimal time step and correction terms (circles) and 
forward Euler with optimal time step but \emph{without} correction terms 
(triangles).  Both schemes use the standard second-order, central finite 
difference approximation for the Laplacian.
Using each method, we solve the diffusion equation with $D = 0.5$ and
source term 
$f(x) = 5 \sin(3 \pi x/2) - 7 \sin(15 \pi x/2) + 10 \sin(21 \pi x/2)$
to time $t = 0.5$ on the domain $0 < x < 1$ subject to the boundary 
conditions $u(0,t) = 1$ and $u_x(1,t) = 1/2$ 
and the initial condition
$u(x,0) = 1 + x/2 + 2 \sin(5 \pi x/2) + 3 \sin(7 \pi x/2) 
- 4 \sin(11 \pi x/2)$.  
These results were obtained using simple MATLAB implementations of the 
finite difference schemes.
}
\label{fig:diffusion_eqn_1d_src_error}
\end{center}
\end{figure}

@@

\begin{figure}[tb]
\begin{center}
\scalebox{0.35}{\includegraphics{figures/adv_diff_eqn_1d_error_vs_N}} 
\caption{$L^\infty$ error as a function of number of grid points for 
two finite difference schemes that solve the advection-diffusion 
equation: forward Euler with optimal time step and correction term (circles) 
and forward Euler with suboptimal time step $\dt = 2\dx^2/3D$ and no correction 
terms (squares).  Both schemes use second-order, central finite difference 
approximations for the advection and diffusion terms.
Using both methods, we solve the advection-diffusion equation with $D = 2$
and $A = 5$ to time $t = 1$ on the domain $-10 < x < 10$ subject to the 
initial condition $u(x,0) = \exp(-x^2/4)$.  The boundary conditions are set 
using the analytical solution 
$u(x,t) = \frac{1}{\sqrt{T}} \exp \left( -(x+At)^2/4T \right)$
where $T = Dt+1$.
These results were obtained using simple MATLAB implementations of the 
finite difference schemes.
}
\label{fig:ADE_1d_error}
\end{center}
\end{figure}

@@

\begin{figure}[tb]
\begin{center}
\scalebox{0.32}{\includegraphics{figures/adv_diff_eqn_1d_FE_OTS_soln}} 
\ \ \ \ \
\scalebox{0.32}{\includegraphics{figures/adv_diff_eqn_1d_FE_soln}} 
\caption{Comparison of numerical solutions for the 1D advection-diffusion
equation~(\ref{eq:ADE_1d}) using forward Euler time stepping with 
optimal time step and correction terms (left) and forward Euler time 
stepping with suboptimal time step $\dt = 2 \dx^2/3 D$ (right).  
Both figures compare the analytical solution (solid line) at $t = 1$ on the 
domain $-10 < x < 10$ with the numerical solution (circles) computed using 
50 grid points.  For details about the initial/boundary 
conditions and physical parameters used to compute these solutions, see the
caption for Figure~\ref{fig:ADE_1d_error}.
}
\label{fig:ADE_1d_solns}
\end{center}
\end{figure}

@@

\subsection{Viscous Burgers Equation}
% FIGURE PLACED HERE TO SPREAD FIGURES OUT
\begin{figure}[tb]
\begin{center}
\scalebox{0.35}{\includegraphics{figures/burgers_eqn_1d_error_vs_N}} 
\caption{$L^\infty$ error as a function of number of grid points for
two finite difference schemes that solve the viscous Burgers equation:
forward Euler with optimal time step and correction term (circles)
and forward Euler with suboptimal time step $\dt = \dx^2/4\nu$ and no 
correction terms (squares).
Both schemes use second-order, central finite difference approximations for
the advection and diffusion terms.
The viscous Burgers equation is solved to time $t=2$ on the domain 
$0 < x < 10$ with the viscosity $\nu$ set to $0.1$.  The initial conditions are 
taken to be 
$u(x,0) = 1 + \gamma \sqrt{\frac{\nu}{\pi}} \exp\left( -(x-1)^2/4 \nu \right) 
            / \left[ 1 + \frac{\gamma}{2} \ 
                         \mathtt{erfc}\left((x-1)/\sqrt{4\nu} \right) 
    \right]$,
where $\ln(1+\gamma)$ is the effective Reynolds number (set equal to 10) and 
the wave speed is set equal to $1$.  The boundary conditions are imposed 
exactly by using
the analytical solution
$u(x,t) = 1 + \gamma \sqrt{\frac{\nu}{\pi T}} 
              \exp\left( -(x-T)^2/4 \nu T \right) 
            / \left[ 1 + \frac{\gamma}{2} \ 
                         \mathtt{erfc}\left((x-T)/\sqrt{4\nu T} \right) 
    \right]$,
where $T = t+1$.
These results were obtained using simple MATLAB implementations of the 
finite difference schemes.
}
\label{fig:burgers_1d_error}
\end{center}
\end{figure}

@@

% FIGURE PLACED HERE TO SPREAD FIGURES OUT
\begin{figure}[tb]
\begin{center}
\scalebox{0.32}{\includegraphics{figures/burgers_eqn_1d_FE_OTS_soln}} 
\ \ \ \ \ 
\scalebox{0.32}{\includegraphics{figures/burgers_eqn_1d_FE_soln}} 
\caption{Comparison of numerical solutions for the 1D viscous Burgers 
equation~(\ref{eq:burgers_1d}) using forward Euler time stepping with 
optimal time step and correction terms (left) and forward Euler time 
stepping with suboptimal time step $\dt = \dx^2/4 \nu$ (right).  
Both figures compare the analytical solution (solid line) at $t = 2$ on the 
domain $0 < x < 10$ with the numerical solution (circles) computed using 
$100$ grid points.  
For details about the initial/boundary conditions and physical parameters 
used to compute these solutions, see the caption for 
Figure~\ref{fig:burgers_1d_error}. 
}
\label{fig:burgers_1d_solns}
\end{center}
\end{figure}

@@

\begin{figure}[tb]
\begin{center}
\scalebox{0.33}{
  \includegraphics{figures/4th_order_parabolic_eqn_1d_error_vs_N}} 
\ \ 
\scalebox{0.33}{
  \includegraphics{figures/4th_order_parabolic_eqn_1d_comp_time}} 
\caption{$L^\infty$ error as a function of number of grid points (left)
and computation time as a function of $L^\infty$ error (right) for two 
finite difference schemes that solve the fourth-order parabolic equation 
(\ref{eq:4th_order_parabolic_eqn_1d_space_err}):
forward Euler with optimal time step and correction term (circles)
and forward Euler with suboptimal time step $\dt = \dx^2/16$ and no 
correction terms (squares).
Both schemes use a fourth-order, central finite difference approximation for
the bilaplacian.  Using each method, we solve the fourth-order parabolic
equation with $\kappa = 1$ to time $t = 0.001$ on the domain $0 < x < 1
$ subject to the boundary conditions $u(0,t) = u(1,t) = 0$ and the initial 
condition $u(x,0) = 2\sin(2 \pi x) + \sin(5 \pi x)$.  The value of the solution
in the ghost cells (required as a result of the wide stencil for the 
discrete bilaplacian) are filled by taking values from the exact solution: 
$u(x,t) = 2\sin(2 \pi x)\cos(10000 \pi t) 
        + \sin(5 \pi x) \left[ 2-\exp(-100\pi^2 t) \right]$.
These results were obtained using simple MATLAB implementations of the 
finite difference schemes.
}
\label{fig:4th_order_parabolic_eqn_1d_error}
\end{center}
\end{figure}

@@

\begin{figure}[tb]
\begin{center}
\scalebox{0.33}{\includegraphics{figures/diffusion_eqn_2d_src_error_vs_N}} 
\ \ 
\scalebox{0.33}{\includegraphics{figures/diffusion_eqn_2d_src_comp_time}} 
\caption{$L^\infty$ error as a function of number of grid points (left)
and computation time as a function of $L^\infty$ error (right) for three
finite difference schemes that solve the 2D diffusion equation:
forward Euler with optimal time step and correction terms (circles), forward 
Euler with suboptimal time step $\dt = \dx^2/4D$ (squares), and 
Crank-Nicholson with time step $\dt = \dx/16D$ (diamonds).  
The forward Euler schemes use a nine-point discretization of the Laplacian;
the Crank-Nicholson scheme uses the standard five-point discretization of
the Laplacian.  Using each method, we solve the 2D diffusion equation with 
$D = 0.25$ and source term 
$f(x,y,t) = 8 D \pi^2 \sin(2 \pi x) \sin(3 \pi y) 
            \exp\left(-5 D \pi^2 t\right)
          - 24 D \pi^2 \sin(5 \pi x) \sin(\pi y) 
            \exp \left(-2 D \pi^2 t \right)$.
to time $t = 0.1$ on the domain $0 < x, y < 1$ subject 
to the initial condition
$u(x,y,0) =  1 + y/3 + x/2 + xy/4 
          + \sin(2 \pi x) \sin(3 \pi y) - \sin(5 \pi x) \sin(\pi y)$.
The boundary conditions are taken from the analytical solution 
$u(x,t) = 1 + y/3 + x/2 + xy/4 
          + \sin(2 \pi x) \sin(3 \pi y) \exp(-5 D \pi^2 t) 
          - \sin(5 \pi x) \sin(\pi y) \exp(-2 D \pi^2 t)
$.
Note that the errors for the forward Euler with suboptimal time step and 
Crank-Nicholson schemes lie almost directly on top of each other at the 
resolution of the figures.  
These results were obtained using simple MATLAB implementations of the 
finite difference schemes.
}
\label{fig:diffusion_eqn_2d_src_analysis}
\end{center}
\end{figure}

@@

% FIGURE SHOWING SOLUTION OF 2D DIFFUSION EQUATION ON STARFISH DOMAIN
% PLACED HERE TO SPREAD FIGURES OUT
\begin{figure}[tb]
\begin{center}
\scalebox{0.35}{\includegraphics{figures/diffusion_eqn_2d_starfish_domain_soln}} 
\scalebox{0.35}{\includegraphics{figures/diffusion_eqn_2d_starfish_domain_error}} 
\caption{Numerical solutions (left) and dominant error (right) for the 2D 
diffusion equation on a starfish-shaped domain.  The solution is computed at 
time $t = 0.1$ using forward Euler time stepping with an optimal time step 
and correction terms.  In the error plot, the dark regions represent points 
where the error in the solution is larger than $25$\% of the $L^\infty$ error 
of the solution.  For these figures, we use the same diffusion constant, 
source term, and initial conditions as described in the caption for 
Figure~\ref{fig:diffusion_eqn_2d_src_analysis}.  The boundary of the 
starfish-shaped domain is represented by the zero-level set of the function
$\sqrt{x^2 + y^2} - \frac{3}{5} \left(1 + \frac{1}{2}\sin(5 \theta) \right)$, 
where $\theta = \mathtt{atan}\left( y/x \right)$.
The Dirichlet boundary conditions are taken from the values of the analytical 
solution (see caption for Figure~\ref{fig:diffusion_eqn_2d_src_analysis})
on the irregular boundary.  The numerical solution is calculated using the 
square computational domain $-1 < x,y < 1$ which is discretized using 400 grid 
points in each coordinate direction (the plot of the solution has been 
downsampled for visualization purposes).  Boundary conditions are imposed by 
using fourth-order accurate approximations of the value of the solution in 
edge and corner ghost cells (described in the main text).
}
\label{fig:diffusion_eqn_2d_starfish_domain}
\end{center}
\end{figure}

@@

% FIGURE SHOWING SOLUTION OF 2D DIFFUSION EQUATION ON STARFISH DOMAIN
% PLACED HERE TO SPREAD FIGURES OUT
\begin{figure}[tb]
\begin{center}
\scalebox{0.35}{
  \includegraphics{figures/diffusion_eqn_2d_starfish_domain_error_vs_N}} 
\caption{$L^\infty$ error as a function of number of grid points for two
finite difference schemes that solve the 2D diffusion equation on an 
irregular domain: forward Euler with optimal time step and correction 
terms (circles) and forward Euler with suboptimal time step $\dt = \dx^2/8 D$
(squares).  Both methods solve the 2D diffusion equation described in the 
caption for Figure~\ref{fig:diffusion_eqn_2d_src_analysis} on the irregular
domain defined in the caption for 
Figure~\ref{fig:diffusion_eqn_2d_starfish_domain}.
These results were obtained using simple MATLAB implementations of the 
finite difference schemes.
}
\label{fig:diffusion_eqn_2d_starfish_error}
\end{center}
\end{figure}
